---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



```{r echo=TRUE, message=FALSE}
library(dplyr)
library(data.table)
```

## Loading and preprocessing the data

```{r echo=TRUE}
if (!file.exists("Data/activity.zip")) {
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    zfile="Data/activity.zip"
    download.file(url, zfile)
    unzip(zipfile=zfile, exdir="Data")
}
```


```{r echo=TRUE}
df <- read.csv("Data/Activity.csv")
df <- mutate(df, id = rownames(df))

```


Ignore Missing Values
```{r echo=TRUE}
noNA <- na.omit(df)

```


##  What is the mean total number of steps taken per day?
```{r echo=TRUE}
noNA <- data.table(noNA, key="date")
dailySteps <- noNA [,sum(steps), by=key(noNA)]
setnames(dailySteps, "V1", "steps")
summary(dailySteps)
```


```{r echo=TRUE}

hist(dailySteps$steps, col="green", 
      main="", xlab="daily steps")
rug(dailySteps$steps)
abline(v=median(dailySteps$steps), lwd=4, col="blue")

```


## What is the average daily activity pattern?

```{r echo=TRUE}

vy <- tapply(noNA$steps, noNA$interval, mean)
vx <- as.integer(names(vy))

plot(vx, vy, type="l", xlab="Intervals", 
     ylab="Average Steps", 
     main="")

```


The 5-minute interval that has the maximum number of steps

```{r echo=TRUE}
z <- as.data.frame(vy)
vy[which(z$vy == max(z$vy))]

```



## Imputing missing values
The toal number of missing values in the dataset. (rows with NA)
```{r echo=TRUE}
nrow(subset(df, is.na(steps) ==TRUE))

```


Fill missing data with imputed mean values 
```{r echo=TRUE,  message=FALSE}

library(Hmisc)
df.i <- with(df, impute(steps, mean))
df$impmean <- as.integer(df.i)

```


There is a little difference in mean and median values.
```{r echo=TRUE}
dt <- data.table(df, key="date")
ds <- dt[,sum(impmean), by=key(dt)]
setnames(ds, "V1", "impmean")
summary(ds)
```



It appears there is little impact of imputing missing data on the estimates of the total daily number of steps.
```{r echo=TRUE}
# png(file="Figures/plot3.png", width=480, height=480)
hist(ds$impmean, col="green",  
     main="Daily Steps Histogram with Imputed Mean Values", xlab="daily steps")
rug(ds$impmean)
abline(v=median(ds$impmean), lwd=4, col="blue")
# dev.off()
```




## Are there differences in activity patterns between weekdays and weekends?

```{r echo=TRUE}
df$isweekend <- weekdays(as.Date(df$date)) %in% c("Saturday", "Sunday")
df$isweekend <- factor(c(TRUE,FALSE))

s <- split(df,df$isweekend)
wkend <-s$'TRUE' %>% filter(!is.na(steps))
wkday <- s$'FALSE' %>% filter(!is.na(steps))

library(ggplot2)
we <- tapply(wkend$steps, wkend$interval, mean)
wd <- tapply(wkday$steps, wkday$interval, mean)

x1 <- as.numeric(names(we))
g1 <- qplot(x1,we, aes(x=interal, y=steps) )
g1 + geom_line()  + coord_cartesian(ylim=c(0,120))  + labs(title="Week End")

```

```{r}
x2 <- as.numeric(names(wd))
g2 <- qplot(x2,wd, aes(x=interal, y=steps) )
g2 + geom_line()  + coord_cartesian(ylim=c(0,120))  + labs(title="Week Days")

```


