# Activity Monitor
### Load and preprocess the data
- Load the data (read.csv())
- Process/transform data (if necessary) into a format suitable for analysis

```{r echo=TRUE}
if (!file.exists("../Data/activity.zip")) {
    url <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
    zfile="../Data/activity.zip"
    download.file(url, zfile)
    unzip(zipfile=zfile, exdir="../Data")
}

library(data.table)

df <- read.csv("../Data/Activity.csv")
dt <- data.table(df, key="date")
dt[,"asDate"] <- as.Date(dt[,date])
dt[,"wday"] <- weekdays(dt$asDate)
head(dt)
```

###### (Reshaped Dataset)


### 1. What is mean total number of steps taken per day?
- Calculate the total number of steps taken per day (ignore the missing values).
- Make a histogram of the total number of steps taken per day.
- Calculate and report the mean and median of the total number of steps taken per day.
```{r}
noNA <- na.omit(dt)
dailySteps <- noNA[, sum(steps), by=key(dt)]
setnames(dailySteps, "V1", "steps")
cat("Mean is ", mean(dailySteps$steps), " Median is: ", median(dailySteps$steps))

cat("Maximum Steps is:", max(dailySteps$steps))
breaks <- c(0,2500,5000,7500,10000,12500,15000,17500,20000,22500)
cuts <- cut(dailySteps$steps, breaks, 
            labels=c("2500", "5000", "7500", "10000", "12500", "15000",
                      "17500",  "20000", "22500"))
summary(cuts)

hist(dailySteps$steps, col="green", 
     breaks=breaks,
     main="Daily Steps Histogram", xlab="daily steps")
rug(dailySteps$steps)
abline(v=median(dailySteps$steps), lwd=4, col="blue")
```



### 2. what is the average daily activity pattern?
- Make a time series plot (type="1") of the 5 minute interval (x-axis) and averaged across all days (y-axis)
```{r echo=TRUE}

vy <- tapply(noNA$steps, noNA$interval, mean)
vx <- as.integer(names(vy))
plot(vx, vy, type="l", xlab="Intervals", 
     ylab="Average Steps", 
     main="Average Daily Activity Pattern")

round(vy,digit=2)
```

- Which 5 minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}
z <- as.data.frame(vy)
vy[which(z$vy == max(z$vy))]

```


### 3. Impute missing values
The presense of missing days may introduce bias
into some calculations or summaries of data.

- Calculate and report the toal number of missing values in the dataset. (rows with NA)
- Devise a strategy for filling in all of the missing values in the dataset
- Create a new dataset dat is equal to the original dataset but with missing data filled in
- Make a historam of the toal number of steps taken each day and calculate the mean and median total number of steps taken per day.  Do these values differ from the estimates from the first part of assignment?  
- What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r echo=TRUE}

```



### 4. Are there differences in activity patterns between weekdays and weekends?
- Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

- Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r echo=TRUE}

```




